---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
## load needed libraries

```{r}
library(plyr)
library(ggplot2)
library(data.table)

```

## unzip activity.zip to read the csv

```{r}
unzip(file.path(getwd(), "activity.zip") ,exdir=".")

```

## Loading and preprocessing the data


```{r}
activity_data <- data.table(read.csv('activity.csv', header = TRUE))
activity_data$DateField <- as.Date(activity_data$date)
na_data <- complete.cases(activity_data)
activity_data1 <- activity_data[na_data,]
summ_per_day <- ddply(activity_data1, "DateField", summarize, 
                      steps=sum(steps),
                      interval=sum(interval) )

```

```{r}
qplot(steps, data=summ_per_day, binwidth=1000)
```

## What is mean total number of steps taken per day?
```{r echo=FALSE}
ototal <- sum(summ_per_day$steps)
omean <- mean(summ_per_day$steps)
omedian <- median(summ_per_day$steps)

```
### Summary (Data With NA)
- Total steps: `r ototal`
- Mean Steps Per Day: `r omean`
- Median of steps Per Day: `r omedian`

## What is the average daily activity pattern?
```{r}
daily_intervals <- activity_data[, list(ms = mean(steps, na.rm=TRUE)), interval]
ggplot(daily_intervals, aes(x=interval, y=ms)) + geom_line()
```

## Imputing missing values
```{r  echo=FALSE}
all_na = sum(is.na(activity_data))
```

There are `r all_na` missing Values

```{r}
imputed_data <- transform(activity_data, steps = ifelse(is.na(activity_data$steps),                                                        daily_intervals$ms[match(activity_data$interval,                                                                       daily_intervals$interval)], 
                                      activity_data$steps))

imp_summ_per_day <- ddply(imputed_data, "DateField", summarize, 
                      steps=sum(steps),
                      interval=sum(interval) )

qplot(steps, data=imp_summ_per_day, binwidth=1000)

```
```{r computetime, echo=FALSE}
itotal <- sum(imp_summ_per_day$steps)
imean <- mean(imp_summ_per_day$steps)
imedian <- median(imp_summ_per_day$steps)

diff_total <- ototal -itotal
diff_mean <- omean -imean
diff_median <- omedian -imedian

```
### Summary (Data after imputing)
- Total steps: `r itotal`
- Mean Steps Per Day: `r imean`
- Median of steps Per Day: `r imedian`

### Difference in Data after imputing
- Total steps: `r diff_total`
- Mean Steps Per Day: `r diff_mean`
- Median of steps Per Day: `r diff_median`


## Are there differences in activity patterns between weekdays and weekends?
```{r}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
imputed_data$dow = as.factor(ifelse(is.element(weekdays(as.Date(imputed_data$DateField)),weekdays), 
                                    "Weekday", "Weekend"))

steps_by_interval_i <- aggregate(steps ~ interval + dow, imputed_data, mean)

library(lattice)

xyplot(steps_by_interval_i$steps ~ steps_by_interval_i$interval|steps_by_interval_i$dow, main="Average Steps per Day by Interval",xlab="Interval", ylab="Steps",layout=c(1,2), type="l")

```